---
title: "302_A3_model_tests_and_selection"
author: "Amanda Ng"
date: "2023-11-22"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
data <- read_csv("cleaned_data.csv")
library(car)
library(MASS)
library(dplyr) 
```

# Assumption and transformation
```{r}
# Original model
model<- lm(`Daily stress` ~ `Social Network` + `Daily Steps` + `Daily Shouting` + `Sleep` + `Time for Passion` + `Meditation` + `Gender`, data)

summary(model)
```

```{r}
# Sleep variance stabalizing transformation
# Make an appropriate plot of Sleep
hist(data$Sleep)
#Left skewed, so we try square
```

```{r}
# transform your response
data$squareSleep <- data$Sleep*data$Sleep

# fit a new model using transformed response
new_model<- lm(`Daily stress` ~ `Social Network` + `Daily Steps` + `Daily Shouting` + squareSleep + `Time for Passion` + `Meditation` + `Gender`, data)
```

```{r}
# Check assumptions and conditions for new model

# Generate all assumption plots for new model
e_hat <- resid(new_model)
y_hat <- fitted(new_model)
# plot residuals vs fitted value plot
par(mfrow=c(1,2))
plot(x = y_hat, y = e_hat, xlab="Fitted Values", ylab="Residuals")

# residuals vs predictors
plot(x = data$`Daily Steps`  , y =  e_hat   , main="Residual vs Social_Network",
     xlab="Social_Network", ylab="Residual")
plot(x = data$`Daily Shouting`    , y =  e_hat   , main="Residual vs DAILY_SHOUTING",
     xlab="DAILY_SHOUTING", ylab="Residual")
plot(x = data$squareSleep    , y =  e_hat   , main="Residual vs Hours Slept",
     xlab="Hours Slept", ylab="Residual")
plot(x = data$`Time for Passion`    , y =  e_hat   , main="Residual vs TIME_FOR_PASSION ",
     xlab="TIME_FOR_PASSION", ylab="Residual")
plot(x = data$`Meditation` , y =  e_hat   , main="Residual vs WEEKLY_MEDITATION ",
     xlab="WEEKLY_MEDITATION", ylab="Residual")

# boxplot of residuals vs Men
boxplot(e_hat ~ data$Gender, main="Residuals by gender", xlab="Gender", ylab="Residuals",
        names=c("Men", "Women"))

# QQ-plot
qqnorm(e_hat)
qqline(e_hat)

# plot to check additional condition 1
plot(y_hat, data$`Daily stress`, main="Daily stress vs Fitted", xlab="Fitted", ylab="Daily stress")

# plot to check additional condition 2
pairwise <- data%>%dplyr::select(`Social Network`, `Daily Steps` , `Daily Shouting`,squareSleep,`Time for Passion`,`Meditation`,`Gender`)

pairs(pairwise[,])
```


# Tests

```{r}
# F test
summary(new_model)
```

Since p-value is small (< 0.05 alpha level), we would reject the null hypothesis and conclude that a significant linear relationship exists between the response and at least one of the predictors.

Based on the summary from the model, we decide to eliminates any predictors that aren’t significantly linearly related to the response (i.e. social network and daily steps) (p-value large)

```{r}
# Reduced model
reduced_model<- lm(`Daily stress` ~ `Daily Shouting` + squareSleep + `Time for Passion` + `Meditation` + `Gender`, data)
```

```{r}
# Check assumptions and conditions for reduced model

# Generate all assumption plots for reduced model
e_hat <- resid(reduced_model)
y_hat <- fitted(reduced_model)
# plot residuals vs fitted value plot
par(mfrow=c(1,2))
plot(x = y_hat, y = e_hat, xlab="Fitted Values", ylab="Residuals")

# residuals vs predictors
plot(x = data$`Daily Steps`  , y =  e_hat   , main="Residual vs Social_Network",
     xlab="Social_Network", ylab="Residual")
plot(x = data$`Daily Shouting`    , y =  e_hat   , main="Residual vs DAILY_SHOUTING",
     xlab="DAILY_SHOUTING", ylab="Residual")
plot(x = data$squareSleep    , y =  e_hat   , main="Residual vs Hours Slept",
     xlab="Hours Slept", ylab="Residual")
plot(x = data$`Time for Passion`    , y =  e_hat   , main="Residual vs TIME_FOR_PASSION ",
     xlab="TIME_FOR_PASSION", ylab="Residual")
plot(x = data$`Meditation` , y =  e_hat   , main="Residual vs WEEKLY_MEDITATION ",
     xlab="WEEKLY_MEDITATION", ylab="Residual")

# boxplot of residuals vs Men
boxplot(e_hat ~ data$Gender, main="Residuals by gender", xlab="Gender", ylab="Residuals",
        names=c("Men", "Women"))

# QQ-plot
qqnorm(e_hat)
qqline(e_hat)

# plot to check additional condition 1
plot(y_hat, data$`Daily stress`, main="Daily stress vs Fitted", xlab="Fitted", ylab="Daily stress")

# plot to check additional condition 2
pairwise2 <- data%>%dplyr::select(`Daily Shouting`,squareSleep,`Time for Passion`,`Meditation`,`Gender`)

pairs(pairwise2[,])
```

```{r}
# Conduct a partial F test
anova(reduced_model, new_model)
```

We fail to reject the null hypothesis and conclude that the predictors we removed from the full model were
all not significantly linearly related to the response.


# Multicollinearity
```{r}
# find the VIF values of all predictors in the model
vif(reduced_model)
```
All well below 5, no multicollinearity problem.

# Problematic points
```{r}
# useful values:
n <- nrow(data)
p <- length(coef(reduced_model))-1
# leverage cutoff
h_cut <- 2*(p+1)/n
# cooks cutoff
D_cut <- qf(0.5, p+1, n-p-1)
# DFFITS cutoff
fits_cut <- 2*sqrt((p+1)/n)
# DFBETAS cutoff
beta_cut <- 2/sqrt(n)


# leverage
h_ii <- hatvalues(reduced_model)
# outlier
r_i <- rstandard(reduced_model)
# Cook's Distance
D_i <- cooks.distance(reduced_model)
# DFFITS
dffits_i <- dffits(reduced_model)
# DFBETAS
dfbetas_i <- dfbetas(reduced_model)
```

```{r}
# identify leverage points
which(h_ii > h_cut)
```

```{r}
# identify outliers
which(r_i > 4 | r_i < -4)
```
No outliers


```{r}
# influential on all fitted values
which(D_i > D_cut)
```
No influential poiunts on all fitted values


```{r}
# influential on own fitted value
which(abs(dffits_i) > fits_cut)
```

```{r}
# influential on a coefficient
for(i in 1:6){
print(paste0("Beta ", i-1))
print(which(abs(dfbetas_i) > beta_cut)) # add your criteria here - this checks all betas in a loop
}
```

# Model selection
```{r}
stepAIC(lm(`Daily stress` ~ `Daily Shouting` + squareSleep + `Time for Passion` + `Meditation` + `Gender`, data),
        direction = "both", k = 2)
```
The full model is the best model.
Full model = sleep transformed and 2 predictors removed from partial F test.